---
title: "Integrate Data - Bcells"
author: Tobias Roider
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y, %X')`"
output: html_document# Find
editor_options: 
  chunk_output_type: console
---

```{r options, include=FALSE, warning = FALSE}
library(knitr)
opts_chunk$set(echo=TRUE, tidy=FALSE, include=TRUE, message=FALSE,
               dpi = 100, cache = FALSE, warning = FALSE)
options(bitmapType = "cairo")

```

# Load packages and functions
```{r Load packages and functions}

library(Seurat)
library(tidyverse)
library(readxl)

source("code/SeuratProc.R")
source("code/QCplots.R")

```

# Read and merge data
```{r}

sobjs_B <- readRDS(file = "output/List_SeuratObjects_B.rds")

# Remove empty Objects
sobjs_B <- compact(sobjs_B)
length(sobjs_B)

# Merge objects
for(i in 1:length(sobjs_B)) {
  if(i==1){
    Combined_B <- merge(sobjs_B[[1]], sobjs_B[[2]])
  }
  if(i>2){
    Combined_B <- merge(Combined_B, sobjs_B[[i]])
  }
}

rm(sobjs_B)

```

```{r eval=T, include=F}

Combined_B <- subset(Combined_B, subset = Entity!="CLL")

```

# Split objects by run
```{r}

splitted_objects <- SplitObject(Combined_B, split.by = "Run")

```

# RNA
## Find anchors and integrate data
```{r}

anchors <- FindIntegrationAnchors(object.list = splitted_objects, 
                                  dims = 1:20, 
                                  assay = rep("RNA", length(splitted_objects)))

Combined_B <- IntegrateData(anchorset = anchors, 
                                new.assay.name = "integratedRNA")

DefaultAssay(Combined_B) <- "integratedRNA"

```

## Revemove cells with missing values
```{r}

cells_remove <- 
  Combined_B@assays$integratedRNA@data %>% 
  as.matrix() %>% 
  is.na() %>% .[1, ] %>% which %>% names

length(cells_remove)

Combined_B <- subset(Combined_B, cells = setdiff(colnames(Combined_B), cells_remove))

```

## Standard workflow for integrated object
```{r}

Combined_B <- ScaleData(Combined_B, verbose = FALSE) 
Combined_B <- RunPCA(Combined_B, 
                     reduction.name = "pcaRNA", reduction.key = "pcaRNA_")

Combined_B <- RunUMAP(Combined_B, dims = 1:20, reduction.key = "umapRNA_",
                      reduction.name = "umapRNA", reduction = "pcaRNA")

Combined_B <- FindNeighbors(Combined_B, reduction = "pcaRNA", dims = 1:20)
Combined_B <- FindClusters(Combined_B, resolution = 0.4)
Combined_B <- FindClusters(Combined_B, resolution = 0.6)
Combined_B <- FindClusters(Combined_B, resolution = 0.8)

```

## Visualization
```{r}

DimPlot(Combined_B, reduction = "umapRNA", label = T)

DimPlot(Combined_B, reduction = "umapRNA", label = F, group.by = "PatientID")+
  theme(legend.position = "none")

```

## Plot heatmap
```{r fig.height=8}

markers <-  FindAllMarkers(Combined_B)
top_markers <- 
  markers %>% 
  group_by(cluster) %>% 
  top_n(5, avg_log2FC) %>% pull(gene)

DoHeatmap(Combined_B, features = top_markers, size=3)

```

# ADT
## Find anchors and integrate data
```{r}

anchors <- FindIntegrationAnchors(object.list = splitted_objects, 
                                  dims = 1:20, 
                                  assay = rep("ADT", length(splitted_objects)))

Combined_B_ADT <- IntegrateData(anchorset = anchors,
                                new.assay.name = "integratedADT")

Combined_B[["integratedADT"]] <- Combined_B_ADT[["integratedADT"]]

rm(Combined_B_ADT)

```

## Revemove cells with missing values
```{r}

cells_remove <- 
  Combined_B@assays$integratedADT@data %>% 
  as.matrix() %>% 
  is.na() %>% .[1, ] %>% which %>% names

length(cells_remove)

Combined_B <- subset(Combined_B, cells = setdiff(colnames(Combined_B), cells_remove))

```

## Standard workflow for integrated object
```{r}

DefaultAssay(Combined_B) <- "integratedADT"

# Run the standard workflow for visualization and clustering
Combined_B <- ScaleData(Combined_B, verbose = FALSE)
Combined_B <- RunPCA(Combined_B, npcs = 30, nfeatures.print = 5,
                         reduction.name = "pcaADT", reduction.key = "pcaADT_")

Combined_B <- RunUMAP(Combined_B, reduction = "pcaADT", dims = 1:20, 
                          reduction.name = "umapADT", 
                          reduction.key = "umapADT_")

Combined_B <- FindNeighbors(Combined_B, reduction = "pcaADT", 
                                dims = 1:20)
Combined_B <- FindClusters(Combined_B, resolution = 0.4)
Combined_B <- FindClusters(Combined_B, resolution = 0.6)
Combined_B <- FindClusters(Combined_B, resolution = 0.8)

```

## Visualization
```{r}

DimPlot(Combined_B, reduction = "umapADT", label = TRUE)
DimPlot(Combined_B, reduction = "umapADT", 
        label = FALSE, group.by = "PatientID")+
  theme(legend.position = "none")

```

## Plot heatmap
```{r}

markers <-  FindAllMarkers(Combined_B)
top_markers <- 
  markers %>% 
  group_by(cluster) %>% 
  top_n(5, avg_log2FC) %>% pull(gene)

DoHeatmap(Combined_B, features = top_markers, size=4)

```

# Save object
```{r}

saveRDS(Combined_B, file = "output/SeuratObjects_Int/Combined_B.rds")

```
