---
title: "Seurat Processing - Run 18"
author: Tobias Roider
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r options, include=FALSE, warning = FALSE}
library(knitr)
opts_chunk$set(echo=TRUE, include=TRUE, message=TRUE,
               dpi = 100, cache = FALSE, warning = FALSE)
options(bitmapType = "cairo")

```

# Load packages and data
```{r message=FALSE}

library("readxl")
library("tidyverse")
library("Seurat")
source("code/SeuratProc.R")
source("code/QCplots.R")
load("output/sampleList.RData")

select <- dplyr::select
filter <- dplyr::filter
mutate <- dplyr::mutate
summarise <- dplyr::summarise

run_current <- "Run 18"

```

# Read samples of this run
```{r}

samples <- sampleList %>% filter(Run==run_current) %>% pull(PatientID)

samples

sobjs <- 
  lapply(samples, function(x){
  z <- list.files(path = "output/SeuratObjects_Raw_5prime", pattern = x, full.names = T)
  y <- readRDS(z)
  return(y)
})

```

# Process samples
## First sample
### Pre-Processing
```{r message=T}

samples[1]
sobj_tmp <- sobjs[[1]]

# Set processing parameters
nFeature_RNA_low=800
percent.mt_top=12
percent.mt_low=0.75
nFeature_RNA_top=6000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Remove cluster 7 (doublets)
```{r}

sobj_tmp <- subset(sobj_tmp, idents = setdiff(c(0:12), 7))
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "2", "3", "7")
Tc <- c("1", "4", "5", "6", "8")
O <-  c("9", "10", "11")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[1]] <- sobj_tmp

```

## Second sample
### Pre-Processing
```{r message=T}

samples[2]
sobj_tmp <- sobjs[[2]]

# Set processing parameters
nFeature_RNA_low=1000
percent.mt_top=10
percent.mt_low=0.5
nFeature_RNA_top=7750
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "2", "5", "7")
Tc <- c("1", "3", "4", "6", "10")
O <-  c("8", "9", "11")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[2]] <- sobj_tmp

```

## Third sample
### Pre-Processing
```{r message=T}

samples[3]
sobj_tmp <- sobjs[[3]]

# Set processing parameters
nFeature_RNA_low=800
percent.mt_top=10
percent.mt_low=0.75
nFeature_RNA_top=6000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("1", "2", "3", "10")
Tc <- c("0", "4", "5", "6")
O <-  c("7", "8", "9")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[3]] <- sobj_tmp

```

## Fourth sample
### Pre-Processing
```{r message=T}

samples[4]
sobj_tmp <- sobjs[[4]]

# Set processing parameters
nFeature_RNA_low=1300
percent.mt_top=8
percent.mt_low=0.75
nFeature_RNA_top=7000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "2", "3", "7", "8", "9")
Tc <- c("1", "4", "5", "6")
O <- c("10", "11")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[4]] <- sobj_tmp

```

# Final objects
```{r}

sobjs

```

# Save objects
```{r}

for(i in 1:length(sobjs)){
  saveRDS(object = sobjs[[i]], file = paste0("output/SeuratObjects_Proc_5prime/SeuratObjectProc_", 
                                             samples[i], ".rds"))
  }

```
