---
title: "Seurat Processing - Run 17"
author: Tobias Roider
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
editor_options: 
  chunk_output_type: console
---

```{r options, include=FALSE, warning = FALSE}
library(knitr)
opts_chunk$set(echo=TRUE, include=TRUE, message=TRUE,
               dpi = 100, cache = FALSE, warning = FALSE)
options(bitmapType = "cairo")

```

# Load packages and data
```{r message=FALSE}

library("readxl")
library("tidyverse")
library("Seurat")
source("code/SeuratProc.R")
source("code/QCplots.R")
load("output/sampleList.RData")

select <- dplyr::select
filter <- dplyr::filter
mutate <- dplyr::mutate
summarise <- dplyr::summarise

run_current <- "Run 17"

```

# Read samples of this run
```{r}

samples <- sampleList %>% filter(Run==run_current) %>% pull(PatientID)

samples

sobjs <- 
  lapply(samples, function(x){
  z <- list.files(path = "output/SeuratObjects_Raw_5prime", pattern = x, full.names = T)
  y <- readRDS(z)
  return(y)
})

```

# Process samples
## First sample
### Pre-Processing
```{r message=T}

samples[1]
sobj_tmp <- sobjs[[1]]

# Set processing parameters
nFeature_RNA_low=800
percent.mt_top=10
percent.mt_low=0.75
nFeature_RNA_top=6000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "1", "4", "7")
Tc <- c("2", "3")
O <-  c("5", "6")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[1]] <- sobj_tmp

```

## Second sample
### Pre-Processing
```{r message=T}

samples[2]
sobj_tmp <- sobjs[[2]]

# Set processing parameters
nFeature_RNA_low=700
percent.mt_top=10
percent.mt_low=0
nFeature_RNA_top=4500
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "3", "10")
Tc <- c("1", "2", "4", "5")
O <-  c("6", "7", "8", "9")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[2]] <- sobj_tmp

```

## Third sample
### Pre-Processing
```{r message=T}

samples[3]
sobj_tmp <- sobjs[[3]]

# Set processing parameters
nFeature_RNA_low=800
percent.mt_top=10
percent.mt_low=0.5
nFeature_RNA_top=6000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "1", "5")
Tc <- c("2", "3", "4", "6")
O <-  c("7")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[3]] <- sobj_tmp

```

## Fourth sample
### Pre-Processing
```{r message=T}

samples[4]
sobj_tmp <- sobjs[[4]]

# Set processing parameters
nFeature_RNA_low=800
percent.mt_top=10
percent.mt_low=0
nFeature_RNA_top=7000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing I
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

# Check out expression of C1Q
FeaturePlot(sobj_tmp, features = "C1QA")

```

### Post-processing II
```{r}

# Remove Cluster 6 (low mito) and 8 (C1QA)
sobj_tmp <- subset(sobj_tmp, idents=c(0:5,7,9))
sobj_tmp <- SeuratProc5prime(sobj_tmp)

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("2", "3", "6", "7")
Tc <- c("0", "1", "4", "5", "8")
O <- ("")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[4]] <- sobj_tmp

```


## Fifth sample
### Pre-Processing
```{r message=T}

samples[5]
sobj_tmp <- sobjs[[5]]

# Set processing parameters
nFeature_RNA_low=900
percent.mt_top=15
percent.mt_low=0
nFeature_RNA_top=6000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing I
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check out expression of C1Q
FeaturePlot(sobj_tmp, features = "C1QA")

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Post-processing II
```{r}

# Remove Cluster 6 (low mito) and 7 (C1QA)
sobj_tmp <- subset(sobj_tmp, idents=c(0:5,8:11))
sobj_tmp <- SeuratProc5prime(sobj_tmp)

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Post-processing III
```{r}

# Remove Cluster 8 (low mito)
sobj_tmp <- subset(sobj_tmp, idents=c(0:7, 9:10))
sobj_tmp <- SeuratProc5prime(sobj_tmp)

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "4", "6", "9")
Tc <- c("1", "2", "3", "5", "7", "8")
O <-  c("10", "11")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    ) %>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[5]] <- sobj_tmp

```

## Sixth sample
### Pre-Processing
```{r message=T}

samples[6]
sobj_tmp <- sobjs[[6]]

# Set processing parameters
nFeature_RNA_low=1000
percent.mt_top=10
percent.mt_low=0.5
nFeature_RNA_top=5000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "4", "5")
Tc <- c("1", "2", "3", "6")
O <-  c("7", "8", "9", "10")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[6]] <- sobj_tmp

```

## Seventh samples
Unsufficient quality (maybe wetting failure)

## Eigth sample
### Pre-Processing
```{r message=T}

samples[8]
sobj_tmp <- sobjs[[8]]

# Set processing parameters
nFeature_RNA_low=850
percent.mt_top=10
percent.mt_low=0.5
nFeature_RNA_top=7000
dims.clustering=1:14
resolution.clustering = 0.4
dims.umap=1:14

# Basic QC plots (pre-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# Run Seurat Processing
sobj_tmp <- SeuratProc5prime(sobj_tmp)

```

### Post-processing I
```{r}

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check out expression of C1Q
FeaturePlot(sobj_tmp, features = "C1QA")

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Post-processing II
```{r}

# Remove Cluster 6 (C1QA)
sobj_tmp <- subset(sobj_tmp, idents=c(0:5, 7:12))
sobj_tmp <- SeuratProc5prime(sobj_tmp)

# Basic QC plots (post-processing)
VlnPlot_custom(sobj_tmp)
FeatureScatter_custom1(sobj_tmp)
FeatureScatter_custom2(sobj_tmp)  

# UMAP Plot
DimPlot_custom1(sobj_tmp)

# Check clustering by features or mito genes
FeaturePlot_custom1(sobj_tmp)

# Check doublets of B and T cells
FeaturePlot_custom3(sobj_tmp)

```

### Assign cell types by RNA cluster
```{r}

B <-  c("0", "1", "8", "9")
Tc <- c("2", "3", "4", "5", "6", "7")
O <-  c("10", "11", "12")

sobj_tmp$CT_rnaCluster <- 
  left_join(
    data.frame(sobj_tmp@meta.data),
    tibble(seurat_clusters=c(B, Tc, O) %>% as.factor(), 
           CT_rnaCluster=c(rep("B", length(B)), 
                           rep("T", length(Tc)), 
                           rep("Other", length(O))) %>% as.factor()
          )
    )%>% pull(CT_rnaCluster)

DimPlot_custom2(sobj_tmp)

sobjs[[8]] <- sobj_tmp

```

# Final objects
```{r}

sobjs

```

# Save objects
```{r}

for(i in 3){
  saveRDS(object = sobjs[[i]], file = paste0("output/SeuratObjects_Proc_5prime/SeuratObjectProc_", 
                                             samples[i], ".rds"))
  }

```
