---
title: 'Figure 6: Intratumor maturation states occupy distinct spatial microenvironments'
author: "Donnacha Fitzgerald"
date: "2023-04-14"
output: html_document
---

This script generates plots in Figure 6. It follows the analyses in CODEX/. See CODEX/CellularNeighborhoods.ipynb for the neighborhood analysis and for spatial plots.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(workflowr)
library(tidyverse)
library(patchwork)
```

```{r}
# Set color palette for maturation states
mathue9 <- c("#ff707c",
              "#975F24",
              "#f8d058",
              "#41d5a8",
              "#006c0b",
              "#902395")
```


# Load Data

```{r}
# Load Seurat object containing CODEX data
CODEX <- readRDS(file = "data/Objects/CODEX_filtered.rds")
```

```{r}
# Set reduction based on x and y coordinates for spatial plots
dimreds <- function(object){
  df <- data.frame(coord_1 = object$x, coord_2 = object$y, row.names = object$unique_cell_id) %>% as.matrix()
  dimred <- CreateDimReducObject(embeddings=df, global = TRUE)
  object@reductions[["coord"]] <- dimred
  return(object)
}

downsampled <- dimreds(downsampled)
CODEX <- dimreds(CODEX)
```

```{r}
# Set cell type factor order
CODEX$CellTypes_final <- factor(CODEX$CellTypes_final, levels = c(
  "B Naïve", "B DZ", "B LZ", "B Mem-IgM", "B Mem-IgG", "B Plasma",
  "CD4T_naive", "TH_memory", "TFH", "TREG", "TPR", 
  "CD8T_naive", "TTOX_memory", "TTOX_exh",
  "NK", "NKT",
  "DC", "FDC",
  "MC", "Macro", "Granulo",
  "Stromal"
  ))
```

```{r}
# Update memory B cell labels consistent with other figures
relabel <- function(obj){
  Idents(obj) <- "predicted.Maturation_course2"
  obj <- RenameIdents(object = obj, 
                    `MD27` = "Mem-IgM",
                    `IgG` = "Mem-IgG")
  obj$B_States <- Idents(obj)
  return(obj)
}
CODEX <- relabel(CODEX)
CODEX$B_States <- factor(CODEX$B_States, levels = c("NA", "Naïve", "DZ", "LZ", "Mem-IgM", "Mem-IgG", "Plasma"))
```

```{r, eval = FALSE}
# Merge unassigned neighborhood
CODEX$N20C11 <- CODEX$N20C11 %>% as.factor()
table(CODEX$N20C11)
Idents(CODEX) <- 'N20C11'
CODEX <- RenameIdents(object = CODEX, `-1` = "0")
CODEX$N20C11 <- Idents(CODEX)
table(CODEX$N20C11)
```

```{r, eval = FALSE}
# Add neighborhood labels
new.cluster.ids <- c("Mem B zone", 
                     "Naïve B enriched",
                     "Plasma/Macro enriched",
                     "Treg/TFH enriched",
                     "Stromal zone",
                     "Follicular",
                     "Mantle zone",
                     "Th Mem zone",
                     "LZ enriched",
                     "Mixed T zone",
                     "DZ enriched")
names(new.cluster.ids) <- levels(CODEX)
CODEX <- RenameIdents(CODEX, new.cluster.ids)
# Order neighborhood labels by B cell subset
Idents(CODEX) <- factor(Idents(CODEX), levels = c(
  "Naïve B enriched",
  "DZ enriched",
  "LZ enriched",
  "Mem B zone",
  "Follicular",
  "Mantle zone",
  "Plasma/Macro enriched",
  "Mixed T zone",
  "Th Mem zone",
  "Treg/TFH enriched",
  "Stromal zone"
))
CODEX$N20C11_labelled <- Idents(CODEX)
```

```{r}
# Save meta data for plotting Figure S6-7.
write_csv(CODEX@meta.data, file = "data/Objects/CODEX_meta.csv", col_names = TRUE)
```

Spatial plots in Figure 6, S6 & S7 were subsequently plotted as in CODEX/CellularNeighborhoods.ipynb.

# CITE-Seq UMAPs
```{r}
Combined <- readRDS("data/Objects/Combined_B_diet.rds")
Combined <- subset(Combined, Entity == "rLN" | Malignancy == "Malignant")
```

```{r}
Samples <- subset(Combined, PatientID == c("LN0259", "LN0135", "LN0014", "LN0188"))
```

```{r, fig.width = 12, fig.height = 3}
# UMAPs
DimPlot(Samples, group.by = "State", split.by = "Entity", cols = mathue9) +
  coord_fixed() +
  ggtitle("Spatial Distribution of B-cell Maturation States") +
  labs(x = "UMAP 1", y = "UMAP 2")+
  theme(axis.text=element_blank(),
        axis.ticks=element_blank()) +
  xlim(c(-6, 9)) +
  ylim(c(-8, 6)) +
  NoLegend() + 
  theme(panel.spacing = unit(3, "cm"))
ggsave("figures/Figure5/UMAPs.pdf")
```

```{r, fig.width = 3, fig.height = 3}
# Accompanying bar charts
ggplot(Samples@meta.data, aes(x = Entity, fill = State)) +
  geom_bar(position = "fill") + 
  ggtitle("") +
  RotatedAxis() +
  theme_nothing() +
  theme(axis.text = element_blank(),
        axis.title = element_blank()) +
  scale_fill_manual(values = mathue9, name = "State") +
  NoLegend()
ggsave("figures/Figure5/MaturationBar.pdf", width = 3, height = 3)
```

# Scipy Interaction Analysis

Spatial graph representations of immediately neighboring cells were constructed based on Delaunay triangulation between centroid coordinates using the scipy.spatial Python package. To compute pairwise association strengths between clusters, relative frequencies were computed using the following metric:

Nij x Nt/
Ni x Nj

in which Nij is equal to the total number of edges between clusters i and j, Nt the total number of edges in the sample, and Ni and Nj the total degrees of clusters i and j respectively.

```{r}
# Import ratio of expected over observed interactions for each cell type pair in each entity (calculated in CODEX/bnhl_graph_analysis.ipynb)
dlbcl_df <- read_csv("/g/huber/projects/CITEseq/CODEX/interaction_analysis/dataframes/dlbcl_df.csv")
fl_df <- read_csv("/g/huber/projects/CITEseq/CODEX/interaction_analysis/dataframes/fl_df.csv")
mcl_df <- read_csv("/g/huber/projects/CITEseq/CODEX/interaction_analysis/dataframes/mcl_df.csv")
mzl_df <- read_csv("/g/huber/projects/CITEseq/CODEX/interaction_analysis/dataframes/mzl_df.csv")
rln_df <- read_csv("/g/huber/projects/CITEseq/CODEX/interaction_analysis/dataframes/rln_df.csv")
```

```{r}
ent_list <- list(dlbcl_df, fl_df, mcl_df, mzl_df, rln_df) %>% lapply(column_to_rownames, "...1")
names(ent_list) <- c("DLBCL", "FL", "MCL", "MZL", "rLN")
ent_list <- ent_list[c("rLN", "MCL", "FL", "DLBCL", "MZL")] # set entity order
ent_list <- lapply(ent_list, function(x) x[-(1:4), 1:4]) # Show only B-cell interactions with other cell types
```

```{r}
# Prepare data for heatmap plotting
prep <- function(df){
  df$row_names <- rownames(df)
  df <- df %>%
  rename_with(~ gsub("B naiive", "B Naïve", .),
              .cols = matches("B naiive")) %>%
  rename_with(~ gsub("B mem", "B Mem", .),
              .cols = matches("B mem")) %>%
  rename_with(~ gsub("plasma", "Plasma", .),
              .cols = matches("plasma"))
  
  # Rename rows
  df <- df %>%
    mutate(
      row_names = case_when(
        row_names == 'B naiive' ~ 'B Naïve',
        row_names == 'B mem' ~ 'B Mem',
        row_names == 'plasma' ~ 'Plasma',
        TRUE ~ row_names
      )
    )
  # Transform the data to long format with 'pivot_longer'
  long_df <- pivot_longer(df, cols = -row_names, names_to = "column_names", values_to = "values")
  
  # Change the order of the factor levels for rows and columns
  long_df$row_names <- fct_relevel(long_df$row_names, "Stromal", "NKT", "NK", "Macro", "MC", "Granulo", "TTOX", "TREG", "TPR", "TH", "TFH", "FDC", "DC", "Plasma", "B Mem", "B GC", "B Naïve")  # Adjust the order as needed
  long_df$column_names <- fct_relevel(long_df$column_names, "B Naïve", "B GC", "B Mem", "Plasma")  # Adjust the order as needed
  
  return(long_df)
}
ent_list_prep <- lapply(ent_list, prep)
```

```{r, fig.width = 10, fig.height = 3}
library(RColorBrewer)
# Create the heatmap
hmap <- function(x){
  plot <- ggplot(x, aes(x = column_names, y = row_names, fill = values)) +
    geom_tile() +
    scale_fill_gradientn(colors = rev(brewer.pal(11, "RdYlBu")), na.value = NA, limits = c(0,1.5)) +
    labs(fill = "IR") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          axis.title = element_blank())
  return(plot)
}
hmap_list <- lapply(ent_list_prep, hmap)
library(patchwork)
patch <- hmap_list$rLN + hmap_list$MCL + hmap_list$FL + hmap_list$DLBCL + hmap_list$MZL + 
  plot_layout(ncol = 5, guides = "collect")
patch
ggsave(plot = patch, path = "figures/Figure5/", filename = "InteractionHeatmaps.pdf", width = 10, height = 3)
```
