---
title: "Fig. 1: A single-cell B-cell maturation reference map in reactive lymph nodes"
author: "Donnacha Fitzgerald"
date: "2023-03-15"
output: html_document
---

This script was used to generate plots for Fig.1. It follows the scripts in the Preprocessing and CelltypeMapping folders.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``` 

```{r}
library(tidyverse)
library(Seurat)
library(SeuratWrappers)
library(questionr)
library(caret)
library(readxl)
library(reader)
library(pheatmap)
```

```{r}
# Set color palette for maturation states
mathue9 <- c("#ff707c", # Naíve
              "#975F24", # DZ
              "#f8d058", # LZ
              "#41d5a8", # Mem IgM
              "#006c0b", # Mem IgG
              "#902395") # Plasma
```

# Load CITE-Seq Data
```{r}
# B cells from reactive lymph node samples only (n=8). See rLN.Rmd for maturation state annotation.
rLN <- readRDS("data/Objects/rLN.rds")
```

See rLN.Rmd for mapping of reactive lymph node B cell maturation states.

# Maturation State UMAP

```{r, fig.width=7, fig.height=7}
DimPlot(rLN,  
        group.by = "State",
        cols = mathue9,
       reduction = "umapRNA",
       shuffle = TRUE) +
  coord_fixed() +
  ggtitle("") +
  labs(x = "UMAP 1", y = "UMAP 2") +
  theme(axis.text=element_blank(),
        axis.ticks=element_blank()) +
  NoLegend()
ggsave("figures/Figure1/MaturationUMAP_coarse.pdf")
```

# Maturation Marker Heatmap
```{r}
# Maturation markers from Supplementary Table 3.
features <- c("IGHD", "IGHM", "TCL1A", "SELL", "CD1C", "CXCR5", 
                         "MME", "BCL6", "CCNB1", "AICDA", 
                         "CAMK1", "CD72", "MS4A1", "PTPN6", 
                         "SLA", "FCRL2", "CFLAR", "FOXP1",
                         "CD83", "EBI3", "BTK", "BLK", "CD74", "BLNK", 
                         "CD40", "NFKB1", "NFKB2", "TRAF1", "ICAM1", "REL", "RELB", "BACH2", 
                         "CCR6", "GPR183", 
                         "TNFRSF17","IRF4", "PRDM1", 
                         "CD38","FAS", "IGHG1", "IGHA1", "IGHE", "CXCR3", 
                         "IGLC2","IGKC")
# Trimmed down to key state-distinguishing features
features3 <- c("IGHD", "SELL",
                "MME", "BCL6", "CCNB1", "AICDA", 
                "CD40", "NFKB2", "CD83", "EBI3", 
                "CCR6", "GPR183",
                "IGHM", "IGHG1", "IGHA1",
               "TNFRSF17","IRF4", "PRDM1")
```

```{r, fig.width = 2.5, fig.height = 4.5}
Idents(rLN) <- "State"
combined_averages <- AverageExpression(rLN, return.seurat = TRUE) 
DoHeatmap(combined_averages, features = features3, label = FALSE, 
          draw.lines = FALSE, group.colors = mathue9, raster = FALSE)   +
    scale_fill_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0) +
  guides(color="none") +
  theme(legend.key.size = unit(0.4, 'cm'), #change legend key size
        legend.text = element_text(size=7),
        legend.title = element_text(size=8)) #change legend title font size 
ggsave(paste0("figures/Figure1/Heatmap_coarse3.pdf"), width = 2.5, height = 4.5)
```

# Confusion Matrix

Generate a confusion matrix of Annotated vs FACS-predicted B cell maturation states.

```{r}
library(cowplot)
conf <- function(model){
  # generate confusion matrix of outer fold predictions vs real classes
  cm <- confusionMatrix(model$State, model$predicted.Maturation.bulk)
  print(cm$overall)
  # plot confusion matrix
  heat_plot = function(x) {
    class(x) = "matrix"
    x = x / rowSums(x) # convert to probability estimates
    pheatmap(x, scale = "none",
             cluster_rows = FALSE, cluster_cols = FALSE,
             color = colorRampPalette(RColorBrewer::brewer.pal(9, "PuBu"))(100),
             breaks = seq(0, 1, by = 0.01)
             )
  }
  plt <- heat_plot(cm$table)
  ggsave(file = "figures/Figure1/CM.pdf", plot = plt, width = 4.5, height = 4)
  plt
}
conf(rLN@meta.data)
```

