---
title: 'Figure S3: B cell maturation marker expression and maturation gene signature scores by entity, related to Figure 2'
author: "Donnacha Fitzgerald"
date: "2023-05-01"
output: html_document
---

This script generates the plots for Figure S3. It follows the analyses in the Preprocessing and CelltypeAnnotation folder.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(Seurat)
library(patchwork)
```

```{r}
# Set color palette for maturation states
mathue9 <- c("#ff707c",
              "#975F24",
              "#f8d058",
              "#41d5a8",
              "#006c0b",
              "#902395")
```

# Malignant B Cell Proportions

```{r}
Combined <- readRDS("data/Objects/Combined_B_full.rds")
Combined <- subset(Combined, subset = Malignancy == "Malignant" | Entity == "rLN")
```

# Maturation Marker Expression

```{r}
# View maturation markers from Morgan ad Tergaonkar et al
features <- c("IGHD", "IGHM", "TCL1A", "SELL", "CD1C", "CXCR5", 
                         "MME", "BCL6", "CCNB1", "AICDA", 
                         "CAMK1", "CD72", "MS4A1", "PTPN6", 
                         "SLA", "FCRL2", "CFLAR", "FOXP1",
                         "CD83", "EBI3", "BTK", "BLK", "CD74", "BLNK", 
                         "CD40", "NFKB1", "NFKB2", "TRAF1", "ICAM1", "REL", "RELB", "BACH2", 
                         "CCR6", "GPR183", 
                         "TNFRSF17","IRF4", "PRDM1", 
                         "CD38","FAS", "IGHG1", "IGHA1", "IGHE", "CXCR3", 
                         "IGLC2","IGKC")
# Revised for the distinction of states
features3 <- c("IGHD", "SELL",
                "MME", "BCL6", "CCNB1", "AICDA", 
                "CD40", "NFKB2", "CD83", "EBI3", 
                "CCR6", "GPR183",
                "IGHM", "IGHG1", "IGHA1",
               "TNFRSF17","IRF4", "PRDM1")
```

```{r}
Entities <- SplitObject(Combined, split.by = "Entity")
```

## Dot Plot

```{r, fig.width = 2.5, fig.height = 4}
matplot <- function(x, group, scale){
  DotPlot(x, assay = "RNA", group.by = group,
            features = rev(features3), dot.scale = scale,
          cols = c("white", "darkblue")) + 
      RotatedAxis() + coord_flip() + ggtitle(x$Entity[1]) +
    NoLegend() +
    theme(axis.title = element_blank())
  ggsave(paste0("figures/Supp5/", x$Entity[1], "DotPlot.pdf"), width = 2.5, height = 4)
}
lapply(Entities, matplot, scale = 4, 
        group = "State")
```

## Gene Signature Scores

Gene signatures for each maturation state were determined from differentially expressed genes in a public tonsil daatset from Holmes et al, JEM, 2020. See CelltypeAnnotation/JEMmaturationscores.Rmd.

```{r}
# Add UPscores as new assay
AddScores <- function(x){
  scores <- colnames(x@meta.data) %>% str_which("UPscore")
  score_matrix <- as.matrix(t(x@meta.data[, scores]))
  rownames(score_matrix) <- gsub(".UPscore", "", rownames(score_matrix))
  x[["scores"]] <-  CreateAssayObject(counts = score_matrix)
  rownames(x[["scores"]]) <- gsub(".UPscore", "", rownames(x[["scores"]]))
  x[["scores"]]@data <- x[["scores"]]@counts
  x <- x %>% ScaleData(assay = "scores")
}
Entities <- lapply(Entities, AddScores)
scores <- rownames(Entities[["DLBCL, non-GCB"]]@assays[["scores"]]@scale.data)
```

```{r, fig.width = 2.5, fig.height = 4}
# Plot scaled scores for each maturation state
ScoresHeatmap <- function(x){
  Idents(x) <- "State"
  DefaultAssay(x) <- "scores"
  AverageExpression(x, return.seurat = TRUE) %>% 
    DoHeatmap(label = TRUE, draw.lines = FALSE, assay = "scores",
              group.colors = mathue9, group.bar = TRUE, size = 4, raster = FALSE,
              features = Features(x)) +
    scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(n =10, name = "RdYlBu")))+
    ggtitle(x$Entity[1]) +
    NoLegend() +
    theme(plot.title = element_text(size=14, vjust = 1, face = "bold"),
          axis.text.y = element_text(size=11))
  ggsave(paste0("figures/Supp5/", x$Entity[1], "ScoresHeatmap.pdf"), width = 2.5, height = 4)
}
lapply(Entities, ScoresHeatmap)
```

# Entity Prediction with Maturation State Composition

```{r}
library(caret)
library(nestedcv)
library(questionr)
library(readxl)
library(reader)
library(pheatmap)
```

```{r, eval = FALSE}
# Calculate the proportions of each maturation state in each sample, and create a data frame for each maturation state.
MatProp <- table(Patient = Combined$PatientID, Entity = Combined$Entity, Maturation = Combined$predicted.Maturation_course) %>% 
  prop.table(margin = 1) %>% 
  as.data.frame() %>% 
  filter(Freq != 0) %>% 
  pivot_wider(names_from = Maturation, values_from = Freq) %>% 
  column_to_rownames(var = "Patient")
# the output is entity class for each sample
Entity <- as.factor(MatProp$Entity)
# the input is maturation state proportion by sample
MatProp <- MatProp[,2:7]
MatProp[is.na(MatProp)] <- 0 # remove NAs
# Define control
trControl <- trainControl(
    number = 10,
    search = "grid")
library(nestedcv)
CITEglmnet <- nestcv.train(Entity,
                                MatProp,
                        trControl = trControl,
                        method = "glmnet",
                        tuneList = NULL,
                        continue_on_fail = TRUE,
                        importance = TRUE,
                        allowParallel = FALSE,
                        metric = "Balanced accuracy",
                        nthread = 1)
saveRDS(CITEglmnet, file = "output/Bcells/FeaturePrediction/BEnt_CITEglmnet.rds")
CITEglmnet$summary

```

```{r, fig.width = 10, fig.height = 4}
CITEglmnet <- readRDS(file = "output/Bcells/FeaturePrediction/BEnt_CITEglmnet.rds")
### Confusion Matrix
cm <- function(model){
  # generate confusion matrix of outer fold predictions vs real classes
  modelcm <- confusionMatrix(model[["output"]][["predy"]], model[["output"]][["testy"]])
  print(modelcm$overall)
  # plot confusion matrix
  heat_plot = function(x) {
    class(x) = "matrix"
    x = x / rowSums(x) # convert to probability estimates
    pheatmap(x, scale = "none",
             cluster_rows = FALSE, cluster_cols = FALSE,
             color = colorRampPalette(RColorBrewer::brewer.pal(9, "PuBu"))(100),
             breaks = seq(0, 1, by = 0.01),
             main = "P( predicted | true )",
             display_numbers = round(x, digits = 2),
             number_color = "orange",
             angle_col = "45")
  }
  confmat <- heat_plot(modelcm$table)
  # plot accuracy statistics
  heat_plot_class = function(x) {
    class(x) = "matrix"
    pheatmap(x, scale = "none",
             cluster_rows = FALSE, cluster_cols = FALSE,
             color = colorRampPalette(RColorBrewer::brewer.pal(9, "PuBu"))(100),
             breaks = seq(0, 1, by = 0.01),
             main = "Test Statistics",
             display_numbers = round(x, digits = 2),
             number_color = "orange",
             angle_col = "45")
  }
  stats <- heat_plot_class(as.matrix(modelcm$byClass))
  patch <- (confmat + stats) +
    plot_layout(guides = "collect")
  patch
}
cm(CITEglmnet)
```

```{r, fig.width = 10, fig.height = 4}
## rf (Random forest)
CITErf <- nestcv.train(Entity,
                        MatProp,
                        trControl = trControl,
                        method = "rf",
                        tuneList = NULL,
                        continue_on_fail = TRUE,
                        importance = TRUE,
                        allowParallel = FALSE,
                        metric = "Balanced accuracy",
                        nthread = 1)
saveRDS(CITErf, file = "output/Bcells/FeaturePrediction/BEnt_CITE_rf.rds")
CITErf$summary
cm(CITErf)
```

```{r, fig.width = 10, fig.height = 4}
## xgbLinear
CITExgbLinear <- nestcv.train(Entity,
                                MatProp,
                        trControl = trControl,
                        method = "xgbLinear",
                        tuneList = NULL,
                        continue_on_fail = TRUE,
                        importance = TRUE,
                        allowParallel = FALSE,
                        metric = "Balanced accuracy",
                        nthread = 1)
saveRDS(CITExgbLinear, file = "output/Bcells/FeaturePrediction/BEnt_CITE_xgbLinear.rds")
CITExgbLinear$summary
cm(CITExgbLinear)
```

```{r, fig.width = 10, fig.height = 4}
## xgbTree
CITExgbTree <- nestcv.train(Entity,
                                MatProp,
                        trControl = trControl,
                        method = "xgbTree",
                        tuneList = NULL,
                        continue_on_fail = TRUE,
                        importance = TRUE,
                        allowParallel = FALSE,
                        metric = "Balanced accuracy",
                        nthread = 1)
saveRDS(CITExgbTree, file = "output/Bcells/FeaturePrediction/BEnt_CITE_xgbTree.rds")
CITExgbTree$summary
cm(CITExgbTree)
```
