---
title: "B Cell Maturation State Label Transfer from CITE-Seq to CODEX"
author: "Donnacha Fitzgerald"
date: "2023-02-28"
output: html_document
---

The purpose of this script is to classify B cell maturation states in CODEX data from CITE-Seq the annotations based on the shared features between both datasets. A logistic regression classifier is trained on the normalized and scaled CITE-Seq features (nested cross-validation), and the resulting best model is applied to the normalized and scaled CODEX features from the isolated B cells. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(23)
```

```{r packages}
library(tidyverse)
library(Seurat)
library(SeuratObject)
library(stringr)
library(caret)
library(questionr)
```

```{r}
# Set color palettes
mathue9 <- c("#ff707c",
              "#975F24",
              "#f8d058",
              "#48edbb",
              "#006c0b",
              "#902395")
```

# Load Data

## CITE-Seq

```{r}
# Load annotated CITE-Seq data from the B cells across all samples.
CITEseq_B <- readRDS("data/Objects/Combined_B_refmap.rds")
DefaultAssay(CITEseq_B) <- "ADT"
Idents(CITEseq_B) <- "predicted.Maturation_course"
DimPlot(CITEseq_B, cols = mathue9, reduction = "ref.umap") + coord_fixed()
```

## CODEX
```{r}
# Load codex data from B cells across all images (after CODEXclustering.Rmd)
CODEX_B <- readRDS("data/Objects/CODEX_B.rds")
```

# Samplewise Classification

```{r}
# Extract sample names
SampleNames <- as.list(unique(CODEX_B$dietrich_id))
# Split the CODEX B cells object by sample
CODEX_B <- SplitObject(CODEX_B, split.by = "dietrich_id")
```

Several machine learning methods, including logistic regression, random forest, gradient boosting decision trees, were first tested on a reactive lymph node sample. As logistic regression showed the highest balanced accuracy in nested cross validation on the CITE-Seq data, and a cleaner visual representation on the lymph node CODEX data, this was pursued on the full dataset.

Random sampling is used to balance classes to improve detection of minority classes (eg. GC or plasma cells).

```{r, results='hide', cache.lazy = FALSE, warning=FALSE, fig.width = 10, fig.height = 8}
# Main function to map maturation state labels data from CITE-Seq to CODEX data
refmap <- function(x){
  # ---- Initialization ----
  gc()  # Run garbage collection to free up memory
  
  # ---- Step 1: Preprocess CITE-Seq Data ----
  # Subset based on specific PatientID
  sample <- subset(CITEseq_B, subset = PatientID == sub('..', '', x))
  
  # Assign cell identities based on maturation state
  Idents(sample) <- sample$predicted.Maturation_course
  
  # Handle singleton cell types
  for (i in levels(sample)){
    if (length(WhichCells(sample, idents = i)) == 1) {
      Idents(sample)[which(Idents(sample) == i)] = "MD27"
      sample$predicted.Maturation_course <- Idents(sample)
    }
  }
  
  # ---- Step 2: Data Scaling ----
  DefaultAssay(sample) <- "codexfeatures"
  sample <- ScaleData(sample)
  
  # Prepare the data matrices
  input <- sample@assays$codexfeatures@scale.data %>% as.matrix() %>% t()
  input[is.na(input)] <- 0
  output <- sample$predicted.Maturation_course %>% as.factor() %>% droplevels()
  
  # ---- Step 3: Model Training ----
  # Set control parameters for model training
  trControl <- trainControl(number = 5, search = "random")
  
  # Seed setting for reproducibility
  set.seed(23)
  
  # Train the glmnet model
  glmnet_sample <- nestcv.train(output, input, 
                                trControl = trControl, 
                                method = "glmnet",
                                tuneList = NULL,
                                continue_on_fail = TRUE,
                                importance = TRUE,
                                allowParallel = FALSE,
                                metric = "Balanced Accuracy",
                                nthread = 1,
                                balance = "randomsample")
  
  # Save the trained model
  saveRDS(glmnet_sample, file = paste0("output/Bcells/CelltypeClassification/Samples/", sample$Entity[1], x, "_Maturation_CODEXfeatures_glmnet.rds"))
  
  # ---- Step 4: Apply Model to CODEX Data ----
  # Extract CODEX data for prediction
  CODEX_B[[x]] <- ScaleData(CODEX_B[[x]])
  codex_counts <- GetAssayData(CODEX_B[[x]], slot = "scale.data")
  # Add . to match with citeseq marker nomenclature
  rownames(codex_counts) <- paste0(".", rownames(codex_counts))
  # Rename features to match citeseq feature names where appropriate
  rownames(codex_counts) <- recode(rownames(codex_counts),
                                   ".lambda" = ".Lambda",
                                   ".kappa" = ".Kappa",
                                   ".CD8" = ".CD8a")
  # subset only features present in CITE-Seq panel
  test <- codex_counts[rownames(codex_counts) %in% ADTfeatures,] %>% as.matrix() %>% t()
  # Predicting on CODEX data
  prediction <- predict(glmnet_sample, newdata = test, verbose = TRUE)
  CODEX_B[[x]]$predicted.Maturation_course <- prediction
  # Plot distribution of predicted maturation states within each entity for CITE-Seq and CODEX
  CODEXplot <- ggplot(CODEX_B[[x]]@meta.data) +
      geom_bar(aes(x = predicted.Maturation_course, fill = predicted.Maturation_course)) +
      RotatedAxis() +
      scale_fill_manual(values = mathue9) +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "CODEX Maturation States")) +
    NoLegend() +
    theme_bw()
  CITEplot <- ggplot(sample@meta.data) +
    geom_bar(aes(x = predicted.Maturation_course, fill = predicted.Maturation_course)) +
      RotatedAxis() +
      scale_fill_manual(values = mathue9) +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "CITE-Seq Maturation States")) +
    NoLegend() +
    theme_bw()
  figure <- CITEplot/CODEXplot
  ggsave(plot = figure, filename = paste0("figures/CelltypeClassification/CODEX/", sample$Entity[1], x, "_MaturationBar_glmnet.pdf"), width = 5, height = 6)
  # Plot spatial distribution
  image1 <- ggplot(CODEX_B[[x]]@meta.data) +
    geom_point(aes(x = Xcorr, y = Ycorr, color = predicted.Maturation_course, alpha = 0.2, stroke = 0.01)) +
    coord_fixed()  +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "Predicted Maturation")) +
      RotatedAxis() +
      scale_color_manual(values = mathue9) +
    theme_classic()
  image2 <- ggplot(CODEX_B[[x]]@meta.data) +
    geom_point(aes(x = Xcorr, y = Ycorr, color = seurat_clusters, alpha = 0.5, stroke = 0.01)) +
    coord_fixed()  +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "Clusters")) +
      RotatedAxis() +
    theme_classic()
  image <- image1/image2
  ggsave(plot = image, filename = paste0("figures/CelltypeClassification/CODEX/", sample$Entity[1], x, "_MaturationImage_glmnet.pdf"), width = 20, height = 15)
  return(CODEX_B[[x]])
}
CODEX_B <- lapply(SampleNames, refmap)
saveRDS(CODEX_B, file = "data/Objects/CODEX_B_refmap.rds")
```

```{r, results='hide', cache.lazy = FALSE, warning=FALSE, fig.width = 10, fig.height = 8}
# Main function to map reference data from CITE-Seq to CODEX data
refmap <- function(x) {
  gc()  # Garbage collection to free up memory

  # ---- Step 1: Preprocess CITE-Seq Data ----
  sample <- subset(citeseq_all, subset = PatientID == sub('..', '', x))
  Idents(sample) <- sample$predicted.Maturation_course  # Set cell identities
  
  # Handle singleton cell types
  for (i in levels(sample)) {
    if (length(WhichCells(sample, idents = i)) == 1) {
      Idents(sample)[which(Idents(sample) == i)] = "MD27"
      sample$predicted.Maturation_course <- Idents(sample)
    }
  }

  # ---- Step 2: Data Scaling ----
  DefaultAssay(sample) <- "codexfeatures"
  sample <- ScaleData(sample)  # Scale data
  input <- sample@assays$codexfeatures@scale.data %>% as.matrix() %>% t()
  input[is.na(input)] <- 0
  output <- sample$predicted.Maturation_course %>% as.factor() %>% droplevels()

  # ---- Step 3: Model Training ----
  trControl <- trainControl(number = 5, search = "random")  # 5-fold cross-validation
  set.seed(23)  # Set random seed for reproducibility
  glmnet_sample <- nestcv.train(output, input, 
                                trControl = trControl, 
                                method = "glmnet", 
                                continue_on_fail = TRUE, 
                                importance = TRUE, 
                                balance = "randomsample",
                                nthread = 1)
  
  # Save the model
  saveRDS(glmnet_sample, file = paste0("output/Bcells/CelltypeClassification/Samples/", sample$Entity[1], x, "_Maturation_CODEXfeatures_glmnet_balanced.rds"))
  glmnet_sample$summary  # Output model summary

  # ---- Step 4: Apply Model to CODEX Data ----
  CODEX_B[[x]] <- ScaleData(CODEX_B[[x]])  # Scale CODEX data
  codex_counts <- GetAssayData(CODEX_B[[x]], slot = "scale.data")

  # Rename features to align with CITE-Seq
  rownames(codex_counts) <- paste0(".", rownames(codex_counts))  # Add prefix "."
  rownames(codex_counts) <- recode(rownames(codex_counts), 
                                   ".lambda" = ".Lambda", 
                                   ".kappa" = ".Kappa", 
                                   ".CD8" = ".CD8a")

  # Subset features present in CITE-Seq
  test <- codex_counts[rownames(codex_counts) %in% ADTfeatures,] %>% as.matrix() %>% t()
  
  # Predict on CODEX data
  prediction <- predict(glmnet_sample, newdata = test, verbose = TRUE)
  CODEX_B[[x]]$predicted.Maturation_course2 <- prediction

  # ---- Step 5: Visualization ----
  # Plot distribution of predicted maturation states within each entity for CITE-Seq and CODEX
  CODEXplot <- ggplot(CODEX_B[[x]]@meta.data) +
      geom_bar(aes(x = predicted.Maturation_course, fill = predicted.Maturation_course)) +
      RotatedAxis() +
      scale_fill_manual(values = mathue9) +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "CODEX Maturation States")) +
    NoLegend() +
    theme_bw()
  CITEplot <- ggplot(sample@meta.data) +
    geom_bar(aes(x = predicted.Maturation_course, fill = predicted.Maturation_course)) +
      RotatedAxis() +
      scale_fill_manual(values = mathue9) +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "CITE-Seq Maturation States")) +
    NoLegend() +
    theme_bw()
  figure <- CITEplot/CODEXplot
  ggsave(plot = figure, filename = paste0("figures/CelltypeClassification/CODEX/", sample$Entity[1], x, "_MaturationBar_glmnet.pdf"), width = 5, height = 6)
  # Plot spatial distribution
  image1 <- ggplot(CODEX_B[[x]]@meta.data) +
    geom_point(aes(x = Xcorr, y = Ycorr, color = predicted.Maturation_course, alpha = 0.2, stroke = 0.01)) +
    coord_fixed()  +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "Predicted Maturation")) +
      RotatedAxis() +
      scale_color_manual(values = mathue9) +
    theme_classic()
  image2 <- ggplot(CODEX_B[[x]]@meta.data) +
    geom_point(aes(x = Xcorr, y = Ycorr, color = seurat_clusters, alpha = 0.5, stroke = 0.01)) +
    coord_fixed()  +
      ggtitle(paste0(CODEX_B[[x]]$entity[1], " ", sub('..', '', CODEX_B[[x]]$dietrich_id[1]), " ", "Clusters")) +
      RotatedAxis() +
    theme_classic()
  image <- image1/image2
  ggsave(plot = image, filename = paste0("figures/CelltypeClassification/CODEX/", sample$Entity[1], x, "_MaturationImage_glmnet.pdf"), width = 20, height = 15)
  return(CODEX_B[[x]])
}
# ---- Main Execution ----
CODEX_B <- lapply(SampleNames, refmap)  # Apply the refmap function to each sample
saveRDS(CODEX_B, file = "data/Objects/CODEX_B_refmap.rds")  # Save processed CODEX data object
```


# Visualizations

## CITE-Seq CODEX Celltype Barcharts

```{r, fig.width = 2, fig.height = 4}
Barplot <- function(x){
  # Plot distribution of predicted maturation states within each entity for CITE-Seq and CODEX
  CODEXplot <- ggplot(x@meta.data) +
      geom_bar(aes(x = predicted.Maturation_course2, fill = predicted.Maturation_course2)) +
      RotatedAxis() +
      scale_fill_manual(values = mathue9) +
      ggtitle(paste0(x$entity[1], " ", sub('..', '', x$dietrich_id[1]), " ", "CODEX")) +
    theme_bw() +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          plot.title = element_text(face="bold", size=12))
  sample <- subset(CITEseq_B, subset = PatientID == sub('..', '', x$dietrich_id[1]))
  CITEplot <- ggplot(sample@meta.data) +
    geom_bar(aes(x = predicted.Maturation_course, fill = predicted.Maturation_course)) +
      RotatedAxis() +
      scale_fill_manual(values = mathue9) +
      ggtitle(paste0(x$entity[1], " ", sub('..', '', x$dietrich_id[1]), " ", "CITE-Seq")) +
    theme_bw() +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          plot.title = element_text(face="bold", size=12))
  CITEplot/CODEXplot
  ggsave(plot = CITEplot/CODEXplot, filename = paste0("figures/CelltypeClassification/CODEX/", x$entity[1], sub('..', '', x$dietrich_id[1]), "_MaturationBar_glmnet_balanced.pdf"), width = 3, height = 6)
}
lapply(CODEX_B, Barplot)
```

## CITE-Seq CODEX celltype proportion correlations

```{r, fig.width=5, fig.height=4}
library(ggpubr)
Matcorr <- function(x){
  sample <- subset(CITEseq_B, subset = PatientID == sub('..', '', x$dietrich_id[1]))
  Idents(sample) <- sample$predicted.Maturation_course
  MatProp <- rbind(table(sample$predicted.Maturation_course)/ncol(sample), table(x$predicted.Maturation_course2)/ncol(x))
  rownames(MatProp) <- c("CITEseq", "CODEX")
  MatProp <- t(MatProp) %>% as.data.frame()
  MatProp$State <- rownames(MatProp) %>% as.factor() %>% factor(levels = c("Naïve", "DZ", "LZ", "MD27", "IgG", "Plasma"))
  ggplot(data = MatProp) +
    geom_point(aes(x = CITEseq, y = CODEX, color = State)) +
    scale_color_manual(values = mathue9) +         
    xlim(0, 1) +
    ylim(0, 1) +
    ggtitle(paste0(x$entity[1], " ", sub('..', '', x$dietrich_id[1]))) +
    theme_bw() +
    theme(legend.position = "none") +
    stat_cor(mapping = aes(x = CITEseq, y = CODEX), p.accuracy = 0.001, r.accuracy = 0.01)
  ggsave(filename = paste0("figures/CelltypeClassification/CODEX/", x$entity[1], sub('..', '', x$dietrich_id[1]), "_MaturationCorr_glmnet_balanced.pdf"), width = 2, height = 2)
}
lapply(CODEX_B, Matcorr)
```
