---
title: "CODEX Analysis with Seurat"
author: "Donnacha Fitzgerald"
date: '2022-11-23'
output: html_document
---

Here, we subset the B cells' data from all CODEX images after segmentation. Normalizatiom, PCA and clustering is performed for exploring the data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
set.seed(23)
```

```{r packages, eval = TRUE}
library(tidyverse)
library(Seurat)
library(SeuratObject)
```

# Load CODEX data

```{r, eval = FALSE}
# Load CODEX data after segmentation and annotation
obj <- readRDS("data/20221006_stanford_seurat_final.rds")
# Filter non-cells
obj <- subset(obj, subset = Merged_final_binned %in% c("noise"), invert = T)
# Bin some annotations
Idents(obj) <- "Merged_final_binned"
obj <- RenameIdents(object = obj, `B_prol` = "B", `Mesenchymal` = "Stromal", `BEC` = "Stromal", `LEC` = "Stromal")
obj$CellTypesDF <- Idents(obj)
```

```{r load}
obj <- readRDS("data/Objects/CODEX.rds")
colnames(obj@meta.data)
```

# Reductions and Clustering
```{r obj-normalization}
# increase maximum array memory size
options(future.globals.maxSize= 8000 * 1024^2)
obj <- NormalizeData(object = obj, normalization.method = "CLR", margin = 2)
obj <- ScaleData(obj)
VariableFeatures(obj) <- rownames(obj)  # since the panel is small, treat all features as variable.
```

```{r obj-pca}
obj <- RunPCA(object = obj, npcs = 20)
```

```{r obj-umap}
obj <- RunUMAP(object = obj, dims = 1:20)
```

```{r obj-neighbors}
obj <- FindNeighbors(object = obj, dims = 1:20)
```

```{r obj-clustering}
obj <- FindClusters(object = obj, resolution = 0.4, n.start = 1)
```

```{r obj-save}
saveRDS(obj, "data/Objects/CODEX.rds")
```


# Split by Sample
```{r, eval = FALSE}
samples <- SplitObject(obj, split.by = "sample_name")
saveRDS(samples, "data/Objects/CODEXlist.rds")
rm(samples)
```

# B Cell Clustering
```{r subset}
# Isolate all B cells (3 million)
Bcells <- subset(obj, idents = c("B", "PC")) %>% RenameAssays(RNA  = 'CODEX')
rm(obj)
```


```{r B-processing}
Bcells <- NormalizeData(object = Bcells, normalization.method = "CLR", margin = 2)
Bcells <- ScaleData(Bcells)
VariableFeatures(Bcells) <- rownames(Bcells)  # since the panel is small, treat all features as variable.
Bcells <- RunPCA(object = Bcells, npcs = 20, verbose = FALSE)
Bcells <- RunUMAP(object = Bcells, dims = 1:20, verbose = FALSE)
gc()
DimPlot(Bcells, group.by = "dietrich_id", shuffle = TRUE)
DimPlot(Bcells, group.by = "entity", shuffle = TRUE)
DimPlot(Bcells, group.by = "sample_name", shuffle = TRUE)
```

```{r}
Bcells <- FindNeighbors(object = Bcells, dims = 1:20, verbose = FALSE)
Bcells <- FindClusters(object = Bcells, verbose = FALSE, resolution = 0.5, n.start = 1)
saveRDS(Bcells, "data/Objects/CODEX_B.rds")
```


```{r}
DimPlot(Bcells, shuffle = TRUE)
DimPlot(Bcells, group.by = "dietrich_id", shuffle = TRUE)
```
CLusters largely reflect individual samples, indicating a strong batch effect.

```{r, fig.width= 7, fig.height = 7}
DoHeatmap(subset(Bcells, downsample = 100), features = VariableFeatures(Bcells))
```
```{r}
FeaturePlot(Bcells, features = c("CD45RA", "BCL6", "CD57", "CD69", "CD11c", "CD38"))
```
```{r}
DotPlot(Bcells, features = c("CD45RA", "BCL6", "CD57", "CD69", "CD11c", "CD38"))
```

## Integration by sample

Perform data integration based on MNN and CCA for batch effect removal. The output is stored in a separate 'integrated' assay.
```{r integration, eval = TRUE}
Bcells <- readRDS("data/Objects/CODEX_B.rds")
# increase maximum array memory size
library(unix)
rlimit_as(1e400)
gc()
Bcells.anchors <- FindIntegrationAnchors(object.list = SplitObject(Bcells, split.by = "dietrich_id"))
gc()
# this command creates an 'integrated' data assay
Bcells <- IntegrateData(anchorset = Bcells.anchors)
gc()
# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(Bcells) <- "integrated"
# Run the standard workflow for visualization and clustering
Bcells <- ScaleData(Bcells, verbose = FALSE)
Bcells <- RunPCA(Bcells, npcs = 20, verbose = FALSE, reduction.name = "pca_int")
Bcells <- RunUMAP(Bcells, reduction = "pca_int", dims = 1:20, reduction.name = "umap_int")
Bcells <- FindNeighbors(Bcells, reduction = "pca_int", dims = 1:20)
Bcells <- FindClusters(Bcells, resolution = 0.5)
saveRDS(Bcells, "data/Objects/CODEX_B.rds")
```

